The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-27 manuscript.Rmd
finish processing section, closes #39

- `RSelenium`
- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `gce_vm_template()`)
  - To enable quick deployments of key R services such as RStudio and Shiny onto cloud virtual machines (VMs), this package utilises Dockerfiles to move the labour of setting up those services from the user to a premade Docker image. For example, by specifying the template `template="rstudio"` in `gce_vm_template()/gce_vm()` an up to date RStudio Service image is launched.  Specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.  
- [`analogsea`](https://github.com/sckott/analogsea) (digital ocean R client)
- [`plumber`](https://www.rplumber.io/docs/hosting.html#docker)
- [production deployment of neural networks](https://github.com/tmobile/r-tensorflow-api) [@jnolis, @nolistic]
- Kubernetes - A popular platform for managing Docker containers is [Kubernetes](https://kubernetes.io/), which is used for a wide variety of containerized applications. It may be your organisation has a Kubernetes cluster already for other applications. Docker containers are used within Kubernetes clusters to hold native code, for which Kubernetes creates a framework around network connections and scaling of resources up and down. An introduction on using [Kubernetes with R is published at this blog post](https://code.markedmondson.me/r-on-kubernetes-serverless-shiny-r-apis-and-scheduled-scripts/).

@@

First, it is possible to **offload complex processing to a server**.
...

Second, when processes can be **run in parallel** for speeding up R code.

For example, [`googleComputeEngineR`](https://CRAN.R-project.org/package=googleComputeEngineR)'s `gce_vm_cluster()` function can create clusters of 2 or more virtual machines, running multi-CPU architectures. Instead of running a local R script with the local CPU and RAM restrictions, the same code can be processed on all CPU threads of the cluster of machines in the cloud, all running in a Docker container with the same R environments. This is achieved through `googleComputeEngineR`'s integration with the R parralisation library [`library(future)` by Henrik Bengtsson]( https://CRAN.R-project.org/package=future). Local R computation can be thrown up to a multi-CPU and VM environment to achieve parrell computation in a few lines of R code in your local session.
  - [some demonstrations are available here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html).
- [Google Cloud Run](https://cloud.run) - This is a CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure. This dispenses with the developer creating a cloud server to run the Docker image on, by abstracting away those servers to a more serverless configuration. Cloud Run lets you run your code on top of a managed or your own Kubernetes cluster running Knative, and can accept any Docker image. The service takes care of network ingress, scaling machines up and down to zero, authentication and authorisation, all features which are non-trivial for a developer to create on their own. This can be used to scale up R code to millions of instances if they need to, with little or no changes to existing code. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses Cloud Run to create a scalable R plumber API.
https://code.markedmondson.me/googleCloudRunner/index.html

[Google Cloud Build](https://cloud.google.com/cloud-build/) and the Google Container Registry are a continuous intergation service respectively image registry that offload building of images to the cloud, while serving the needs of commercial environments  that helps move the workload of building the Docker images to an online service, while  Cloud Build can be set up to build the Dockerfiles on each GitHub commit or release. This means you do not need to build the Docker images locally, which can tie up resources since Docker images can be several GBs and take a long time to compile. Google Cloud Build works alongside  to allow you to build private and public Docker images, which allows you to build up your own dependency graphs for downstream applications.
  - https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html

- \CRANpkg{batchtools} \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g., start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to Docker&nbsp;Hub if the integration test succeeds.
- \CRANpkg{drake}, https://docs.ropensci.org/drake/index.html?q=docker#with-docker

In the RStudio Server Pro 1.2 release in 2019, RStudio added new functionality called [Launcher](https://solutions.rstudio.com/launcher/overview/).
It gives users the ability to spawn R sessions and background jobs in a scalable way on external clusters, e.g., [Kubernetes based on Docker images](https://support.rstudio.com/hc/en-us/articles/360019253393-Using-Docker-images-with-RStudio-Server-Pro-Launcher-and-Kubernetes) or [Slurm](https://slurm.schedmd.com/) clusters, and optionally, with Singularity containers.
A key benefit of Launcher is the ability for R and Python users to interactively work in RStudio without learning about Docker at all - while still leveraging containers and Kubernetes.
Users can submit batch jobs to Kubernetes clusters without writing specific deployment scripts.

\label{pipelines}
Third, containers are perfectly suited for **packaging and executing software pipelines** and required data.
Containers allow building complex processing pipelines that are independent from the host programming language.
Due to its original use case (see&nbsp;\nameref{introduction}), Docker has no standard mechanisms for chaining containers together; it lacks definitions and protocols for environment variables, volume mounts and/or ports that could enable transfer of input (parameters and data) and output (results) to and from containers.
Some packages, e.g., \pkg{containerit}, do provide Docker images that can be used similarly to CLIs, but their usage is cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.
**\pkg{outsider}** ([https://docs.ropensci.org/outsider/](https://docs.ropensci.org/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own R-based workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

@@

\citet{gentleman_statistical_2007} coined the term **Research Compendium** for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.
In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
Future researchers can download the compendium from the repository and run the included `Dockerfile` to build a new image that recreates the computational environment used to produce the original research results.
If building the image fails, the human-readable instructions in a `Dockerfile` are the starting point for rebuilding the environment.
Further safeguarding practices are currently under development or not part of common practice yet, such as preservation of images \citep{emsley_framework_2018} and storing them alongside `Dockerfile`s \citep[cf.][]{nust_opening_2017}, or pinning of system libraries beyond the tagged base images, which may be seen as stable or dynamic depending on the applied time scale \citep[see discussion on `debian:testing` base image in][]{RJ-2017-065}.

@@

Fourth, containers are useful for **testing** R code during development.
To submit a package to CRAN, an R package must work with the development version of R, which must be compiled locally.
That can be a challenge for some users.
The R-hub service (see Section&nbsp;\nameref{rhub}) makes it easy to ensure that no errors occur, but to fix errors a local setup is still often warranted, e.g., using the image `rocker/r-devel`, and to test packages with native code, which can make the process more complex \citep[cf.][]{eckert_building_2018}.
The R-hub Docker images can also be used to debug problems locally using various combinations of Linux platforms, R versions, and compilers\footnote{See \href{https://r-hub.github.io/rhub/articles/local-debugging.html}{https://r-hub.github.io/rhub/articles/local-debugging.html} and \href{https://blog.r-hub.io/2019/04/25/r-devel-linux-x86-64-debian-clang/}{https://blog.r-hub.io/2019/04/25/r-devel-linux-x86-64-debian-clang/}}.
The images go beyond the configurations, or _flavours_, used by CRAN for checking packages\footnote{\href{https://cran.r-project.org/web/checks/check_flavors.html}{https://cran.r-project.org/web/checks/check\_flavors.html}}, e.g., with CentOS-based images, but lack a container for checking on Windows or OS X.
The images greatly support package developers to provide support on operating systems they are not familiar.
The package \pkg{dockertest} ([https://github.com/traitecoevo/dockertest/](https://github.com/traitecoevo/dockertest/)) is a proof of concept for automatically generating `Dockerfile`s and building images specifically to run tests\footnote{\pkg{dockertest} is not actively maintained, but mentioned still because of its interesting approach.}.
These images are accompanied with a special launch script so the tested source code is not stored in the image but the currently checked in version from a local Git repository is cloned into the container at runtime.
This approach clearly seperates test environment, test code, and current working copy of the code.

@@

Nonetheless, at least on the level of R packages some **consolidation** seems in order, e.g., to reduce the number of packages creating `Dockerfile`s from R code or controlling the Docker daemon with R code.
It remains to be seen which approach to control Docker, via the Docker API as \pkg{stevedore} or via system calls as \pkg{dockyard}/\pkg{docker}/\pkg{dockr}, is more sustainable, or if the question will be answered by the endurance of maintainers and sufficient funding.
Similarly, the capturing of environments and their serialization in form of a `Dockerfile` currently happens at different levels of abstraction and re-use of functionality seems reasonable, e.g., \pkg{liftr} could generate the environment with \pkg{containerit}, which in turn may use \pkg{dockerfiler} for low level R objects representing a `Dockerfile` and its instructions.
In this consolidation, the Rocker&nbsp;Project could play the role of coordinating entity.
Though for the moment, the sign of the times points to more experimentation and feature growth, e.g., images for GPU-based computing and artificial intelligence.
Even with coding being more and more accepted as a required, and achievable skill, an easier access, for example by exposing containerisation benefits via simple user interfaces in the users' IDE, could be an important next step.
Currently containerisation happens more in the background at the system level.

>>>>

- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `gce_vm_template()`)
  - To enable quick deployments of key R services such as RStudio and Shiny onto cloud virtual machines (VMs), this package utilises Dockerfiles to move the labour of setting up those services from the user to a premade Docker image. For example, by specifying the template `template="rstudio"` in `gce_vm_template()/gce_vm()` an up to date RStudio Service image is launched.  Specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.  
- [`analogsea`](https://github.com/sckott/analogsea) (digital ocean R client)
- [`plumber`](https://www.rplumber.io/docs/hosting.html#docker)
- [production deployment of neural networks](https://github.com/tmobile/r-tensorflow-api) [@jnolis, @nolistic]
- Kubernetes - A popular platform for managing Docker containers is [Kubernetes](https://kubernetes.io/), which is used for a wide variety of containerized applications. It may be your organisation has a Kubernetes cluster already for other applications. Docker containers are used within Kubernetes clusters to hold native code, for which Kubernetes creates a framework around network connections and scaling of resources up and down. An introduction on using [Kubernetes with R is published at this blog post](https://code.markedmondson.me/r-on-kubernetes-serverless-shiny-r-apis-and-scheduled-scripts/).

@@

First, it is possible to **offload complex processing to a server** or clouds, and also parallel executing of processes for speeding up or serving of many users.
**\CRANpkg{batchtools}** provides a parallel implementation of Map for various schedulers \citep{Lang2017batchtools}.
The package can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
**\CRANpkg{googleComputeEngineR}** has the function `gce_vm_cluster()` to create clusters of 2 or more virtual machines, running multi-CPU architectures.
Instead of running a local R script with the local CPU and RAM restrictions, the same code can be processed on all CPU threads of the cluster of machines in the cloud, all running a Docker container with the same R environments.
\CRANpkg{googleComputeEngineR} integrates with the R parralisation package \CRANpkg{future} to enable this with only a few lines or R code\footnote{\href{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}}.
[Google Cloud Run](https://cloud.run) is a CaaS (Containers as a Service) platform.
Users can launch containers using any Docker image without worrying about underlying infrastructure in a so called serverless configuration.
The service takes care of network ingress, scaling machines up and down, authentication and authorisation---all features which are non-trivial for a developer to create on their own.
This can be used to scale up R code to millions of instances if need be with little or no changes to existing code, as demonstrated by the proof of concept `cloudRunR` ([https://github.com/MarkEdmondson1234/cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR)), which uses Cloud Run to create a scalable R-based API using \CRANpkg{plumber}.
[Google Cloud Build](https://cloud.google.com/cloud-build/) and the Google Container Registry are a continuous integration service respectively image registry that offload building of images to the cloud, while serving the needs of commercial environments such as private Docker images or image stacks.
As Googl Cloud Build itself can run any container, the package \pkg{googleCloudRunner} demonstrates how R can be used as the control language for one time or batch processing jobs and scheduling of jobs\footnote{\href{https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html}{https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html}}.
\label{drake} **\CRANpkg{drake}** is a workflow manager for data science projects \citep{landau_drake_2019}.
It features implicit parallel computing and automated detection of the parts of the work that actually needs to be reexecuted.
\CRANpkg{drake} has been demonstrated to run inside containers for high reproducibility\footnote{See for example \href{https://github.com/joelnitta/pleurosoriopsis}{https://github.com/joelnitta/pleurosoriopsis} or \href{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}, the latter even running in a CI, see~\nameref{cicd}.}, but also how to use \CRANpkg{future} package's function `makeClusterPSOCK()` to send parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.
In the latter case, the container control code must be written by the user and the \CRANpkg{future} package ensures the host and worker can connect for communicating over socket connections.
**RStudio Server Pro** includes a functionality called [Launcher](https://solutions.rstudio.com/launcher/overview/) (since version&nbsp;1.2, released in 2019).
It gives users the ability to spawn R sessions and background/batch jobs in a scalable way on external clusters, e.g., [Kubernetes based on Docker images](https://support.rstudio.com/hc/en-us/articles/360019253393-Using-Docker-images-with-RStudio-Server-Pro-Launcher-and-Kubernetes) or [Slurm](https://slurm.schedmd.com/) clusters, and optionally, with Singularity containers.
A benefit of the proprietary Launcher software is the ability for R and Python users to leverage containerisation's advantages in RStudio without writing specific deployment scripts or learning about Docker or managing clusters at all.

\label{pipelines}
Second, containers are perfectly suited for **packaging and executing software pipelines** and required data.
Containers allow building complex processing pipelines that are independent from the host programming language.
Due to its original use case (see&nbsp;\nameref{introduction}), Docker has no standard mechanisms for chaining containers together; it lacks definitions and protocols for environment variables, volume mounts and/or ports that could enable transfer of input (parameters and data) and output (results) to and from containers.
Some packages, e.g., \pkg{containerit}, do provide Docker images that can be used similarly to CLIs, but their usage is cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.
**\pkg{outsider}** ([https://docs.ropensci.org/outsider/](https://docs.ropensci.org/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own R-based workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

@@

\citet{gentleman_statistical_2007} coined the term **Research Compendium** for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.
In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
A more extreme case is the _dynverse_ project ([https://dynverse.org/](https://dynverse.org/)), which packages over 50 computational methods with different environments (R, Python, C++, etc.) in Docker images, which can be executed from R.
Future researchers can download the compendium from the repository and run the included `Dockerfile` to build a new image that recreates the computational environment used to produce the original research results.
If building the image fails, the human-readable instructions in a `Dockerfile` are the starting point for rebuilding the environment.
Further safeguarding practices are currently under development or not part of common practice yet, such as preservation of images \citep{emsley_framework_2018} and storing them alongside `Dockerfile`s \citep[cf.][]{nust_opening_2017}, or pinning of system libraries beyond the tagged base images, which may be seen as stable or dynamic depending on the applied time scale \citep[see discussion on `debian:testing` base image in][]{RJ-2017-065}.

@@

Fourth, containers are useful for **testing** R code during development.
To submit a package to CRAN, an R package must work with the development version of R, which must be compiled locally.
That can be a challenge for some users.
The R-hub service (see Section&nbsp;\nameref{rhub}) makes it easy to ensure that no errors occur, but to fix errors a local setup is still often warranted, e.g., using the image `rocker/r-devel`, and to test packages with native code, which can make the process more complex \citep[cf.][]{eckert_building_2018}.
The R-hub Docker images can also be used to debug problems locally using various combinations of Linux platforms, R versions, and compilers\footnote{See \href{https://r-hub.github.io/rhub/articles/local-debugging.html}{https://r-hub.github.io/rhub/articles/local-debugging.html} and \href{https://blog.r-hub.io/2019/04/25/r-devel-linux-x86-64-debian-clang/}{https://blog.r-hub.io/2019/04/25/r-devel-linux-x86-64-debian-clang/}}.
The images go beyond the configurations, or _flavours_, used by CRAN for checking packages\footnote{\href{https://cran.r-project.org/web/checks/check_flavors.html}{https://cran.r-project.org/web/checks/check\_flavors.html}}, e.g., with CentOS-based images, but lack a container for checking on Windows or OS X.
The images greatly support package developers to provide support on operating systems they are not familiar.
The package \pkg{dockertest} ([https://github.com/traitecoevo/dockertest/](https://github.com/traitecoevo/dockertest/)) is a proof of concept for automatically generating `Dockerfile`s and building images specifically to run tests\footnote{\pkg{dockertest} is not actively maintained, but mentioned still because of its interesting approach.}.
These images are accompanied with a special launch script so the tested source code is not stored in the image but the currently checked in version from a local Git repository is cloned into the container at runtime.
This approach clearly seperates test environment, test code, and current working copy of the code.
\label{rselenium} Another use case where a container helps to standardise tests across operating systems is detailed the vignettes of the package \CRANpkg{RSelenium} \citep{rselenium_2019}.
The package recommends Docker for running the [Selenium](https://selenium.dev/) Server application needed to execute test suites on browser-based user interfaces and webpages, but requires users to manually manage the Docker containers.

@@

Nonetheless, at least on the level of R packages some **consolidation** seems in order, e.g., to reduce the number of packages creating `Dockerfile`s from R code or controlling the Docker daemon with R code.
It remains to be seen which approach to control Docker, via the Docker API as \pkg{stevedore} or via system calls as \pkg{dockyard}/\pkg{docker}/\pkg{dockr}, is more sustainable, or if the question will be answered by the endurance of maintainers and sufficient funding.
Similarly, the capturing of environments and their serialization in form of a `Dockerfile` currently happens at different levels of abstraction and re-use of functionality seems reasonable, e.g., \pkg{liftr} could generate the environment with \pkg{containerit}, which in turn may use \pkg{dockerfiler} for low level R objects representing a `Dockerfile` and its instructions.
In this consolidation, the Rocker&nbsp;Project could play the role of coordinating entity.
Though for the moment, the sign of the times points to more experimentation and feature growth, e.g., images for GPU-based computing and artificial intelligence.
Even with coding being more and more accepted as a required, and achievable skill, an easier access, for example by exposing containerisation benefits via simple user interfaces in the users' IDE, could be an important next step.
Currently containerisation happens more in the background at the system level.
Furthermore, a maturing of the Rockerverse packages for managing containers may lead to their adoption where currently manual coding is required, e.g. in the case of \CRANpkg{RSelentium} or \CRANpkg{drake} (see Sections&nbsp;\nameref{rselenium} and \nameref{drake} respectively).
